Transmembrane helices are also an overlooked source of major histocompatibility complex class II epitopes and evolutionary more conserved than expected by chance

### 2021-08-18 bbbq_article.tex
Process feedback, thanks @fransbianchi

[SOMETHING HERE] A peptide is called a binder if, for a certain haplotype, 
any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
all peptides within a \emph{proteome} 
(see supplementary Tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
whereas the previous study defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
This revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes.
To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials.

@@

For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}. Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. This strengthens our previous finding \cite{bianchi2017} and indicates that TMH-derived peptides are over-presented because the peptide binding grooves of most MHC-I and -II haplotypes are relatively hydrophobic.

@@

  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_bw.png}
    \label{fig:f_snps_found_and_expected}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
    \label{fig:conservation}
  \end{subfigure}

@@

  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
    \label{fig:f_tmh_ncbi}
  \end{subfigure}

  \caption{ \textbf{Evolutionary conservation of human TMHs.}
    \textbf{(A)} 
    Percentage of SNPs found in TMHs.
    Each point shows for one protein the predicted percentage of
    TMH ($x$-axis) and the observed occurrence of SNPs being located
    within a TMH ($y$-axis).
    The dashed diagonal line shows the line of equality (i.e.,
    equal conservation of TMHs and soluble protein regions). 
    The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
    \textbf{(B)}
    The number of SNPs in TMHs as expected by chance (left bar) 
    and found in the dbSNP database (right bar).
    Percentages show the relative conservation
    of SNPs in TMHs found.
    \textbf{(C)}
    Distribution of the percentages of TMH in the TMPs used in this study.
  }
\end{figure}

>>>>

We define a protein to be a binder if, for a certain haplotype, 
any of its 9-mer or 14-mer peptides have an IC50 value in the lowest 2\% of 
all peptides within a \emph{proteome} 
(see supplementary Tables \ref{tab:ic50_binders_mhc1} and \ref{tab:ic50_binders_mhc2} for values), 
this differs from our previous study where we defined
a binder as having an IC50 in the lowest 2\% 
of the peptides within a \emph{protein}.
% See https://github.com/richelbilderbeek/bianchi_et_al_2017/blob/72e6755a31d400158368509fd80a41e984677ab1/predict-binders.R#L17
This revised definition precludes bias of proteins 
that give rise to no or only very few MHC epitopes.
To verify that the results are similar, a side by side comparison was performed shown in the supplementary materials.

@@

For MHC-I, we previously showed that the over-presentation of TMH-derived peptides is caused by the hydrophobicity of the peptide binding grooves \cite{bianchi2017}. Figures \ref{fig:hydrophobicity_1} and \ref{fig:hydrophobicity_2}
show the extent of over-presentation
of TMH-derived epitopes as a function of the hydrophobicity preference score for the different haplotypes.
An assumed linear correlation explains 88\% of the variability in MHC-I.
For MHC-II, 62\% of the variability is explained by hydrophobicity. 
This indicates that TMH-derived peptides are over-presented, 
because the peptide binding grooves of most MHC-I and -II haplotypes 
are relatively hydrophobic.

@@

  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_snps_found_and_expected_bw.png}
    \label{fig:f_snps_found_and_expected}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_f_tmh_ncbi.png}
    \label{fig:f_tmh_ncbi}
  \end{subfigure}

@@

  \begin{subfigure}[t]{0.45\textwidth}
    \centering
    \caption{}
    \includegraphics[width=\linewidth]{ncbi_peregrine_results/fig_conservation.png}
    \label{fig:conservation}
  \end{subfigure}

  \caption{ \textbf{Evolutionary conservation of human TMHs.}
    \textbf{(A)} 
    Percentage of SNPs found in TMHs.
    Each point shows for one protein the predicted percentage of
    TMH ($x$-axis) and the observed occurrence of SNPs being located
    within a TMH ($y$-axis).
    The dashed diagonal line shows the line of equality (i.e.,
    equal conservation of TMHs and soluble protein regions). 
    The diagonal line indicates a linear fit, the gray area its 95\% confidence interval.
    \textbf{(B)}
    Distribution of the percentages of TMH in the TMPs used in this study.
    \textbf{(C)}
    The number of SNPs in TMHs as expected by chance (left bar) 
    and found in the dbSNP database (right bar).
    Percentages show the relative conservation
    of SNPs in TMHs found.
  }
\end{figure}
